---
title: "Plant families and orders eaten"
output: html_document
---


```{r global_options, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = TRUE)

```
### Below is a dynamic plot showing the plant families we've eaten. You can hover over the tree tips to see notes. You can collapse or expand individual nodes by clicking on them.

### This figure requires `ate_it.csv`, a file that has a list of species eaten, and any associated notes you'd like to include (e.g., hover over green cabbage.)

```{r}

wd <- "C:/Users/John/Desktop/plant families eaten/"
setwd(wd)

library(ggtree)
library(dplyr)
library(phytools)
library(tidytree)
library(collapsibleTree)
library(data.tree)

phylo <- read.tree('plant_families_Gastauer2017.new')
ate_raw <- read.csv('ate_it.csv', stringsAsFactors = FALSE)
ate <- ate_raw %>%
  mutate_all(trimws) %>%
  mutate(family = tolower(family), 
         species = tolower(species),
         )

```

```{r, fig.height = 10, fig.width = 10, message = FALSE, echo = FALSE}

fams <- unique(ate$family[ate$family != ''])
x <- as_data_frame(phylo)

if(!all(fams %in% unique(ate$all_families))) 
  stop(paste("Not all families eaten are present in tree. Check your spelling of:", fams[!fams %in% unique(ate$all_families)]))
ate <- ate %>%
  filter(family %in% fams) %>%
  mutate(label = paste0(genus, ' ', species, ' (', common_name, ')'))

x$parent_name <- x$label[match(x$parent, x$node)]
x$parent_name[x$label == 'angiosperms'] <- NA

a <- list()
ups <- for(i in fams) {
  #location
  k <-1
  #name
  j <- x$parent_name[x$label == i]
  a[[i]] <- c()
  while (!is.na(j)) {
    a[[i]][[k]] <- j
    k <- k + 1
    j <- x$parent_name[x$label == j]
  }
  a[[i]] <- a[[i]][!grepl("_to_", a[[i]], fixed = TRUE)]
}

j <- data.frame(parent = NA, child = 'angiosperms')

for(i in names(a)) {
  len <- length(a[[i]])
  tmp <- data.frame(parent = a[[i]],
                    child = c(i, a[[i]][1:len-1]))
  j <- rbind(j, tmp)
}

j <- rbind(
  mutate(j,
    color = ifelse(child %in% ate_raw$all_families, as.character(child), NA),
    notes = ''),
  transmute(ate, 
    color = family, 
    parent = family, 
    child = label,
    notes = notes)) %>%
  mutate(
    color = as.numeric(as.factor(color)),
    color = ifelse(is.na(color), '#FFFFFF',
                   colorspace::rainbow_hcl(length(unique(color[!is.na(color)])))[color]))

collapsibleTreeNetwork(j, 
  attribute = 'notes', fill = 'color', 
  collapsed = FALSE, zoomable = FALSE, nodeSize = 'leafCount')

```

### Here is a static plot showing all the families consumed so far.

```{r}

zoom(phylo, fams)

```

### Here is the same thing, but grouped by orders:

```{r, include = FALSE}
ord.tree <- as.Node(phylo)
Prune(ord.tree, function(x) !grepl("eae$", x$levelName))
ord.tree <- ToNewick(ord.tree, heightAttribute = NULL)
ord.tree <- read.tree(text = ord.tree)

```

```{r, fig.height = 10, fig.width = 10, message = FALSE, echo = FALSE}

ords <- unique(ate$order[ate$order != ''])
x <- as_data_frame(ord.tree)

if(!all(ords %in% unique(ate_raw$all_orders))) 
  stop(paste("Not all orders eaten are present in tree. Check your spelling of:", ords[!ords %in% unique(ate$all_orders)]))
ate <- ate %>%
  mutate(label2 = paste0(family, ' ', genus, ' ', species, ' (', common_name, ')'))

x$parent_name <- x$label[match(x$parent, x$node)]
x$parent_name[x$label == 'angiosperms'] <- NA

a <- list()
ups <- for(i in ords) {
  #location
  k <-1
  #name
  j <- x$parent_name[x$label == i]
  a[[i]] <- c()
  while (!is.na(j)) {
    a[[i]][[k]] <- j
    k <- k + 1
    j <- x$parent_name[x$label == j]
  }
  a[[i]] <- a[[i]][!grepl("_to_", a[[i]], fixed = TRUE)]
}

j <- data.frame(parent = NA, child = 'angiosperms')

for(i in names(a)) {
  len <- length(a[[i]])
  tmp <- data.frame(parent = a[[i]],
                    child = c(i, a[[i]][1:len-1]))
  j <- rbind(j, tmp)
}

j <- rbind(
  mutate(j,
    color = ifelse(child %in% ate_raw$all_orders, as.character(child), NA),
    notes = ''),
  transmute(ate, 
    color = order, 
    parent = order, 
    child = label2,
    notes = notes)) %>%
  mutate(
    color = as.numeric(as.factor(color)),
    color = ifelse(is.na(color), '#FFFFFF',
                   colorspace::rainbow_hcl(length(unique(color[!is.na(color)])))[color]))

collapsibleTreeNetwork(j, 
  attribute = 'notes', fill = 'color', 
  collapsed = FALSE, zoomable = FALSE, nodeSize = 'leafCount'
  )


```


### Here is a static plot showing all the orders consumed so far.

```{r}

zoom(ord.tree, ords)

```

### and here are all the orders

```{r, fig.width = 9, fig.height = 7}

ggtree(ord.tree, layout = 'circular') + 
  geom_tiplab2(size = 3)

```

#### These figures are based on the plant family phylogeny from: *Gastauer, Neto, and Alves.* Updated angiosperm family tree for analyzing phylogenetic diversity and community structure. *Acta Botanica Brasilica 31.2 (2017): 191-198.*
